Approximation as Differentiation

### 2025-04-18 appendix/notes/cmon-enriched.tex
Cleanup of enriched functors and enriched Yoneda embedding (when enriching category is concrete).

\subsubsection{$\CMon$-enriched Yoneda embedding}

Suppose $\cat{C}$ a small category. The usual \emph{Yoneda embedding} for $\cat{C}$ is the functor $\Yoneda:
\cat{C} \to \Func{\cat{C}^\op}{\Set}$ defined as follows:
\begin{itemize}
\item on objects: $\Yoneda(X) = \Hom{\cat{C}}{-}{X}: \cat{C}^\op \to \Set$, the contravariant hom-functor
sending:
   \begin{itemize}
   \item any object $Y$ in $\cat{C}$ to the hom-set $\Hom{\cat{C}}{Y}{X}$;
   \item any morphism $f: Y \to Z$ in $\cat{C}$ to the function $\Hom{\cat{C}}{f}{X} = (- \comp f):
   \Hom{\cat{C}}{Z}{X} \to \Hom{\cat{C}}{Y}{X}$.
   \end{itemize}
\item on morphisms: for any $f: X \to Y$ in $\cat{C}$, $\Yoneda(f)$ is the natural transformation where
$\Yoneda(f)_Z = (f \comp -): \Hom{\cat{C}}{Z}{X} \to \Hom{\cat{C}}{Z}{Y}$.
\end{itemize}

\noindent Now suppose $\cat{C}$ is $\mathscr{V}$-enriched with $\mathscr{V}$ monoidal closed; then the
hom-functor $\Hom{\cat{C}}{-}{X}: \cat{C}^\op \to \mathscr{V}$ is itself $\mathscr{V}$-enriched, sending
objects $Y$ to the hom-object $\Hom{\cat{C}}{Y}{X}$ in $\mathscr{V}$, and providing a family of morphisms
$\Hom{\cat{C}}{-}{X}_{Y,Z} \to \Hom{\mathscr{V}}{\Hom{\cat{C}}{Y}{X}}{\Hom{\cat{C}}{Z}{X}}$ for any $X, Y$ in
$\cat{C}$. The following diagram commutes because precomposition with the identity is the identity on
morphisms:

\begin{center}
\begin{tikzcd}[column sep=1.8cm]
I \arrow[d, equals] \arrow[r, "\id_X"] & \Hom{\cat{C}}{X}{X} \arrow[d, "\Hom{\cat{C}}{-}{W}_{X,X}"] \\
I \arrow[r, "\id_{\Hom{\cat{C}}{X}{W}}"'] & \Hom{\mathscr{V}}{\Hom{\cat{C}}{X}{W}}{\Hom{\cat{C}}{X}{W}}\
\end{tikzcd}
\end{center}

\noindent and the following commutes by the associativity of composition:

\begin{center}
\begin{tikzcd}[column sep=2.7cm]
   \Hom{\cat{C}}{Z}{Y} \tensor \Hom{\cat{C}}{Y}{X} \arrow[r, "\comp_{X,Y,Z}"] \arrow[d,
   "\Hom{\cat{C}}{-}{W}_{Y,Z}\,\tensor\,\Hom{\cat{C}}{-}{W}_{X,Y}"'] & \Hom{\cat{C}}{Z}{X} \arrow[d,
   "\Hom{\cat{C}}{-}{W}_{X,Z}"] \\
   \Hom{\mathscr{V}}{\Hom{\cat{C}}{Y}{W}}{\Hom{\cat{C}}{Z}{W}} \tensor
   \Hom{\mathscr{V}}{\Hom{\cat{C}}{X}{W}}{\Hom{\cat{C}}{Y}{W}} \arrow[r,
   "\comp_{\Hom{\cat{C}}{X}{W},\Hom{\cat{C}}{Y}{W},\Hom{\cat{C}}{Z}{W}}"'] &
   \Hom{\mathscr{V}}{\Hom{\cat{C}}{X}{W}}{\Hom{\cat{C}}{Z}{W}}
\end{tikzcd}
\end{center}

\noindent There exists an enriched version of the Yoneda embedding, namely a functor $\Yoneda_{\mathscr{V}}:
\cat{C} \to \Func{\cat{C}^\op}{\mathscr{V}}_{\mathscr{V}}$, which is itself $\mathscr{V}$-enriched, sending
objects $X$ to the $\mathscr{V}$-enriched hom-functor $\Yoneda_{\mathscr{V}}(X) = \Hom{\cat{C}}{-}{X}:
\cat{C}^\op \to \mathscr{V}$ and morphisms $f: X \to Y$ in $\cat{C}$ to $\mathscr{V}$-natural transformations
$\Yoneda_{\mathscr{V}}(f): \Hom{\cat{C}}{-}{X} \naturalto \Hom{\cat{C}}{-}{Y}$ acting at each component as
post-composition with $f$.

The enriched Yoneda lemma says that for any $\mathscr{V}$-enriched presheaf $F$ and any object $X \in
\cat{C}$, the hom-object of $\mathscr{V}$-natural transformations (which also lives in $\mathscr{V}$) between
the $\mathscr{V}$-enriched presheaf $\Hom{\cat{C}}{-}{X}$ and $F$ is isomorphic to $F(X)$:

@@

>>>>

\subsubsection{Enriched Yoneda embedding}

Suppose a monoidal category $\mathscr{V}$ and a $\mathscr{V}$-enriched category $\cat{C}$, again restricted to
the case where $\mathscr{V}$ is concrete.

\begin{definition}[Hom-functor]
Define the contravariant \emph{hom-functor} $\Hom{\cat{C}}{-}{X}: \cat{C}^\op \to \mathscr{V}$ sending
\begin{itemize}
\item any object $Y$ in $\cat{C}$ to the hom-object $\Hom{\cat{C}}{Y}{X}$ in $\mathscr{V}$;
\item any morphism $f: Y \to Z$ in $\cat{C}$ to the function $\Hom{\cat{C}}{f}{X} = (- \comp f):
\Hom{\cat{C}}{Z}{X} \to \Hom{\cat{C}}{Y}{X}$ on the underlying hom-sets.
\end{itemize}
\end{definition}

\begin{definition}[Yoneda embedding]
The \emph{Yoneda embedding} for $\cat{C}$ is the functor $\Yoneda: \cat{C} \to
\Func{\cat{C}^\op}{\mathscr{V}}$ sending
\begin{itemize}
\item any object $X$ to the hom-functor $\Hom{\cat{C}}{-}{X}: \cat{C}^\op \to \mathscr{V}$;
\item any morphism $f: X \to Y$ to the $\mathscr{V}$-natural transformation $\Hom{\cat{C}}{-}{X} \naturalto
\Hom{\cat{C}}{-}{Y}$ where $\Yoneda(f)_Z = (f \comp -): \Hom{\cat{C}}{Z}{X} \to \Hom{\cat{C}}{Z}{Y}$.
\end{itemize}
\end{definition}

\noindent Now suppose $\mathscr{V}$ is monoidal closed. Then the hom-functor $\Hom{\cat{C}}{-}{X}: \cat{C}^\op
\to \mathscr{V}$ is itself $\mathscr{V}$-enriched, providing a family of morphisms $\Hom{\cat{C}}{-}{X}_{Y,Z}:
\Hom{\cat{C}}{Y}{Z} \to \Hom{\mathscr{V}}{\Hom{\cat{C}}{Z}{X}}{\Hom{\cat{C}}{Y}{X}}$ in $\mathscr{V}$ for any
$Y, Z$ in $\cat{C}$. Moreover there exists an enriched Yoneda embedding, namely a functor
$\Yoneda_{\mathscr{V}}: \cat{C} \to \Func{\cat{C}^\op}{\mathscr{V}}_{\mathscr{V}}$ which is itself
$\mathscr{V}$-enriched, sending objects $X$ to the $\mathscr{V}$-enriched hom-functor $\Hom{\cat{C}}{-}{X}$
and morphisms $f: X \to Y$ in $\cat{C}$ to $\mathscr{V}$-natural transformations $\Yoneda_{\mathscr{V}}(f):
\Hom{\cat{C}}{-}{X} \naturalto \Hom{\cat{C}}{-}{Y}$ acting at each component as post-composition with $f$.

The \emph{enriched Yoneda lemma} states that, for any $\mathscr{V}$-enriched presheaf $F$ and any object $X
\in \cat{C}$, there is an isomorphism between $F(X)$ and the hom-object (also living in $\mathscr{V}$) of
$\mathscr{V}$-natural transformations from the $\mathscr{V}$-enriched presheaf $\Hom{\cat{C}}{-}{X}$ to $F$:

@@

% The following diagram commutes because precomposition with the identity is the identity on morphisms:
%
% \begin{center}
% \begin{tikzcd}[column sep=1.8cm]
% I \arrow[d, equals] \arrow[r, "\id_X"] & \Hom{\cat{C}}{X}{X} \arrow[d, "\Hom{\cat{C}}{-}{W}_{X,X}"] \\
% I \arrow[r, "\id_{\Hom{\cat{C}}{X}{W}}"'] & \Hom{\mathscr{V}}{\Hom{\cat{C}}{X}{W}}{\Hom{\cat{C}}{X}{W}}\
% \end{tikzcd}
% \end{center}
%
% \noindent and the following commutes by the associativity of composition:
%
% \begin{center}
% \begin{tikzcd}[column sep=2.7cm]
%    \Hom{\cat{C}}{Z}{Y} \tensor \Hom{\cat{C}}{Y}{X} \arrow[r, "\comp_{X,Y,Z}"] \arrow[d,
%    "\Hom{\cat{C}}{-}{W}_{Y,Z}\,\tensor\,\Hom{\cat{C}}{-}{W}_{X,Y}"'] & \Hom{\cat{C}}{Z}{X} \arrow[d,
%    "\Hom{\cat{C}}{-}{W}_{X,Z}"] \\
%    \Hom{\mathscr{V}}{\Hom{\cat{C}}{Y}{W}}{\Hom{\cat{C}}{Z}{W}} \tensor
%    \Hom{\mathscr{V}}{\Hom{\cat{C}}{X}{W}}{\Hom{\cat{C}}{Y}{W}} \arrow[r,
%    "\comp_{\Hom{\cat{C}}{X}{W},\Hom{\cat{C}}{Y}{W},\Hom{\cat{C}}{Z}{W}}"'] &
%    \Hom{\mathscr{V}}{\Hom{\cat{C}}{X}{W}}{\Hom{\cat{C}}{Z}{W}}
% \end{tikzcd}
% \end{center}
